TITLE by JINSOO KIM
========================================================

<font size=4>

This data set contains 113,937 loans with 81 variables on each loan, including loan amount, borrower rate (or interest rate), current loan status, borrower income, borrower employment status, borrower credit history, and the latest payment information.

Ask your own questions about this data set. There are MANY variables in this data set and you are not expected to explore all of them. You should explore between 10-15 variables in your analysis.



```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(RColorBrewer)
library(tidyverse)


# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

```


```{r echo=FALSE, Load_the_Data}
# Load the Data

#Warning message:
#In strsplit(code, "\n", fixed = TRUE) :
#  input string 1 is invalid in this locale'''

df <- read.csv("C:/Users/Waker/Google 드라이브(1)/Class matrials/DAND2/7. Exploratory Data Analysis (24h) Due. 7 Mar/Project/Data/prosperLoanData.csv")

```

Data wrangling.
Before Analysis I found 2 things to wrangle.
1. Duplicated data
```{r}
count(df) # number of rows in dataset
count(df[unique(df$ListingKey), ]) # number of rows in ListingKey which is unique key for each listing, same value as the 'key' used in the listing object in the API.

```

These are one example of duplicated ListingKey.
```{r}
# Create duplicated dataset
rm_keys <- count(df, ListingKey, sort = T) %>% 
  filter(n > 1)

# Change the name of column
colnames(rm_keys)[2] <- "Number of duplication" 

# Print one example of duplicated
df[(df$ListingKey == "17A93590655669644DB4C06"), ]

```
I found that only the ProsperScore is different. Therefore I will remove five

```{r}
df_no_dupicated <- unique(df[df$ListingKey, ])


df_no_dupicated[(df_no_dupicated$ListingKey == "17A93590655669644DB4C06"), ]
```
힣 모르겠다 어떻게 listingkey 하나만 남기지. 남는 건 어떤 게 남겨지는걸까.



```{r}
df_copy <- df # Copy the original.
```




> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

# Univariate Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

돈 빌려 본 사람 %
```{r}
nrow(subset(df, !is.na(df$TotalProsperPaymentsBilled))) / nrow(df)
```


빌려 본 사람 중 잘 갚은 사람

df_copy <- df_copy %>% 
  filter(!is.na(TotalProsperPaymentsBilled)) %>% 
  mutate(RightTimePayments = (OnTimeProsperPayments + ProsperPaymentsLessThanOneMonthLate) / TotalProsperPaymentsBilled) 

iris$Regulation <- ifelse(iris$Sepal.Length >=5, "UP", "DOWN")
밑에 스택오버플로우에 질문해둠. 새 칼럼 이상하게 생성돼서
```{r}

df_copy$RightTimePayments <- ifelse(df_copy$TotalProsperPaymentsBilled == NA,서 NA, (OnTimeProsperPayments + ProsperPaymentsLessThanOneMonthLate) / TotalProsperPaymentsBilled)

summary(df$RightTimePayments)

```


Proportion of 'Ontime Prosper payments' and 'Less than one month late'
```{r}
df_copy <- df_copy %>% 
  filter(!is.na(TotalProsperPaymentsBilled)) %>% 
  mutate(RightTimePayments = (OnTimeProsperPayments + ProsperPaymentsLessThanOneMonthLate) / TotalProsperPaymentsBilled) %>% 



```


본인이 말한 급여에 따른 아직 안 갚은 돈
```{r echo=FALSE, Univariate_Plots_first}
ggplot(aes(x = StatedMonthlyIncome, y = AmountDelinquent), data = subset(df_copy, !is.na(AmountDelinquent))) +
  geom_point() +
  scale_x_continuous(
    lim = c(0, quantile(df_copy$StatedMonthlyIncome, 0.99))) +
  scale_y_continuous(lim = c(0, quantile(df_copy$AmountDelinquent, 0.99)))
```

***


1. Let's see the distribution of Income Range 
```{r echo=FALSE, Univariate_Plots}
table(df_copy$IncomeRange)
df_copy$IncomeRange <- factor(df_copy$IncomeRange,
                              levels = c('$0', '$1-24,999', '$25,000-49,999', '$50,000-74,999', '$75,000-99,999', '$100,000+', 'Not employed', 'Not displayed'))
ggplot(aes(x = IncomeRange), data = df_copy) +
  geom_bar()
```

2. Occupation distribution
```{r echo=FALSE, Univariate_Plots}

df_copy <- within(df_copy,
                  Occupation <- factor(Occupation, levels = names(sort(table(Occupation)))))
ggplot(aes(Occupation), data = df_copy) +
  geom_bar() +
  coord_flip() +
  xlab('Purpose') +
  ylab('count') +
  ggtitle("Purpose of borrowing")

```

3. Prosper score distribution for 84,853 (29,084 people have no record)
```{r echo=FALSE, Univariate_Plots}
ggplot(df_copy, aes(ProsperScore)) +
  geom_histogram()
```


4. Reasons to borrow money
```{r echo=FALSE, Univariate_Plots}
# Change the number to borrowing purpose
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 0] <- "Not Available"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 1] <- "Debt Consolidation"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 2] <- "Home Improvement"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 3] <- "Business"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 4] <- "Personal Loan"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 5] <- "Student Use"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 6] <- "Auto"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 7] <- "Other"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 8] <- "Baby&Adoption"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 9] <- "Boat"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 10] <- "Cosmetic Procedur"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 11] <- "Engagement Ring"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 12] <- "Green Loans"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 13] <- "Household Expenses"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 14] <- "Large Purchases"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 15] <- "Medical/Dental"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 16] <- "Motorcycle"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 17] <- "RV"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 18] <- "Taxes"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 19] <- "Vacation"
df_copy$ListingCategory..numeric.[df_copy$ListingCategory..numeric. == 20] <- "Wedding Loans"

df_copy <- within(df_copy,
                  ListingCategory..numeric. <- factor(ListingCategory..numeric., levels = names(sort(table(ListingCategory..numeric.)))))
ggplot(aes(ListingCategory..numeric.), data = df_copy) +
  geom_bar() +
  coord_flip() +
  xlab('Purpose') +
  ylab('count') +
  ggtitle("Purpose of borrowing")

```

5. Loan distribtion by states
```{r echo=FALSE, Univariate_Plots}
df_copy <- within(df_copy,
                  BorrowerState <- factor(BorrowerState, levels = names(sort(table(BorrowerState)))))
ggplot(aes(BorrowerState), data = df_copy) +
  geom_bar() +
  coord_flip() +
  xlab('State') +
  ylab('count') +
  ggtitle("Loan distribtion by states")

```

6. LenderYield distribution
```{r echo=FALSE, Univariate_Plots}
ggplot(df_copy, aes(LenderYield)) +
  geom_bar()
```


7. 
```{r echo=FALSE, Univariate_Plots}

```

8. 
```{r echo=FALSE, Univariate_Plots}

```




신용점수 평균내서 보자. 이것도 NA 때문에 못 하고 있음.


> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

This is the 'Top 10 Occupations that borrow the most' except 'Other' and 'blank'.
```{r echo=FALSE, Bivariate_Plots_first}
new <- df_copy %>% 

  group_by(Occupation) %>% 
  summarise(how_many = n()) %>% 
  arrange(-how_many)

occupation_top10_frequent <- head(new, 12)
occupation_top10_frequent <- occupation_top10_frequent[-c(1, 8), ]
ggplot(aes(x = reorder(Occupation, -how_many), y = how_many), data = occupation_top10_frequent) +
  geom_point() +
  xlab('Occupation')
  ylab('Count') +
  ggtitle("Top 10 Occupations that borrow the most except 'Other' and 'blank'")
```

lender yield
This is the Top 10 Occupations that has high lender yield except 'Other' and 'blank'.
```{r echo=FALSE, Bivariate_Plots_second}
High_LenderYield_Occupation <- df_copy %>% 
  filter(LenderYield >= median(df_copy$LenderYield)) %>% 
  group_by(Occupation) %>% 
  summarise(how_many = n()) %>% 
  arrange(-how_many) %>% 
  head(12) %>% 
  
ggplot(aes(x = reorder(Occupation, -how_many), y = how_many), data = High_LenderYield_Occupation[-c(1, 8), ]) +
  geom_point() +
  xlab('Occupation') +
  ylab('count') +
  ggtitle("Top 10 Occupations that has High LenderYiled except 'Other' and 'blank'")

```
"랜더일드 높은 직업"과 "가장 많이 돈 빌린 직업" 9개가 일치. Sales - Retail에 대해 더 공격적으로 돈을 빌려줘야 함!

```{r}
Loan_Purpose <- df_copy %>% 
  group_by(ListingCategory..numeric.) %>% 
  summarise(how_many = n()) %>% 
  arrange(-how_many) %>% 
  head(12)
  

Loan_Purpose$ListingCategory..numeric. <- c('Debt Consolidation', 'Not Available', 'Other', 'Home Improvement', 'Business', 'Auto', 'Personal Loan', 'Household Expenses', 'Medical/Dental', 'Taxes', 'Large Purchases', 'Wedding Loans')

ggplot(aes(ListingCategory..numeric., how_many), data = Loan_Purpose[-c(1, 8), ]) +
  geom_point() +
  xlab('Occupation') +
  ylab('count') +
  ggtitle("Top 10 Occupations that has High LenderYiled except 'Other' and 'blank'")
```

```{r echo=FALSE, Univariate_Plots_seond}
df_occupation_group <- df_copy %>% 
  group_by(Occupation) %>% 
  summarise(how_many = n()) %>% 
  arrange(-how_many) %>% 
  head(12)

ggplot(aes(Occupation, how_many),
       data = df_occupation_group[-c(1, 8), ]) +
  geom_col() +
  xlab('Occupation') +
  ylab('Count') +
  ggtitle("Top 10 Occupations that borrow the most except 'Other' and 'blank'")

```

***



# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!





















</font>